Irreducible Components and Dimension of the Springer Fiber of a Hook-Type Slodowy Slice

### 2024-08-30 paper.tex
thoughts from meeting

\usepackage{hyperref}
\usepackage[capitalize,nameinlink,noabbrev]{cleveref} % to emulate \autoref style

@@

\author{
UROP+ Final Paper, Summer 2024\\
Owen Conoly\\
Mentor: Haoshuo Fu\\
Project Suggested by: TODO: also Haoshuo?
}
\date{}
\title{Title}

\begin{document}

\maketitle
\begin{abstract}

\end{abstract}

\section{Introduction}
Let \(\mathfrak{g} = \gl_m \times \gl_{m + n}\), with \(n > 0\).
% The nilpotent cone in \(\mathfrak{g}\) is \(\mathcal{N} = \mathcal{N}_m \times \mathcal{N}_{m + n}\), where \(\mathcal{N}_m\) and \(\mathcal{N}_{m + n}\) are the nilpotent cones in \(\gl_m\) and \(\gl_{m + n}\) respectively.
% Let \(\mathcal{B}_m\) and \(\mathcal{B}_{m + n}\) be the varieties of Borel subalgebras of \(\gl_m\) and \(\gl_n\) respectively.
% The variety of Borel subalgebras of \(\mathcal{N}\) is \(\mathcal{B} = \mathcal{B}_m \times \mathcal{B}_{m + n}\).
% Let
% \[\widetilde{\mathcal{N}} := \{(\mathfrak{b}, n) \in \mathcal{B} \times \mathcal{N} : n \in \mathfrak{b}\} = \]
% \[\{(\mathfrak{b}_m, \mathfrak{b}_{m + n}, n_m, n_{m + n}) \in \mathcal{B}_m \times \mathcal{B}_{m + n} \times \mathcal{N}_m \times \mathcal{N}_{m + n} : n_m \in \mathfrak{b}_m, n_{m + n} \in \mathfrak{b}_{m + n}\}.\]
Let \(\pi : \widetilde{\mathcal{N}} \to \mathcal{N} \hookrightarrow \mathfrak{g}\) be the Springer resolution.
\par Let \((e, h, f)\) be the principal \(\spl_2\)-triple in \(\gl_n\).
We have an embedding \(\gl_n \hookrightarrow \gl_m \times \gl_n \hookrightarrow \gl_{m + n}\).
Let \((E, H, F)\) be the \(\spl_2\)-triple in \(\gl_{m + n}\) which is the image of \((e, h, f)\) via this embedding.
Let \(S\) be the Slodowy slice \(E + \mathfrak{z}_{\gl_m}(F)\).
We have a map \(\gl_{m + n} \to \gl_m\) given by coordinate projection.
Restricting to \(S\), we obtain \(f_1 : S \to \gl_m\).
(It turns out that this map is surjective.)
Then, we obtain \(p_1 : S \to \mathfrak{g}\) by \(x \mapsto (x, f_1(x))\).
Taking the map \(p_1 : S \to \mathfrak{g}\) and the Springer resolution \(\widetilde{\mathcal{N}} \to \mathfrak{g}\), we obtain a fibred product \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).
\par In this paper we study \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).
% The first few sections will be finding what it looks like (in particular, what are the elements of \(S\) which project to nilpotent elements of \(\mathfrak{g}\)), and then we will find the irreducible components of \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).
As a step towards studying \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\), we consider the map \(S \times_\mathfrak{g} \widetilde{\mathcal{N}} \to \gl_m\) given by taking the map to \(\mathfrak{g}\) and then projecting to \(\gl_m\).
The main section of this paper will study the fibers of this map.
Given \(X \in \mathcal{N}_m \subseteq \gl_m\), we call the fiber at \(X\) the \emph{\(n\)-Slodowy-slice Springer fiber at \(X\)}.
\par Section 2 reviews some preliminary material.
Section 3 finds the unique (up to similarity) principal \(\spl_2\)-triple in \(\gl_n\).
Section 4 embeds this \(\spl_2\)-triple into \(\gl_{m + n}\) as described above.
We compute the Slodowy slice and end up with a nice description of \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).
\par Section 5 discusses how to reduce the problem of finding the irreducible components of an \(n\)-Slodowy-slice Springer fiber at \(X \in \mathcal{N}_m\) to an easier problem.
Section 6 solves the easier problem.
Section 7 finds the irreducible components of an \(n\)-Slodowy-slice Springer fiber.
Section 8 applies the results of section 7 to find the irreducible components of \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).
Section 9 discusses an interesting direction for future work.
Finally, section 10 proves some linear algebra lemmas that were used in the paper.

@@

\subsection{Springer fibers in \texorpdfstring{\(\gl_m\)}{gl\_m}}
Now we let \(\mathcal{N}_m\) be the set of nilpotent elements in \(\gl_m\), and \(\mathcal{B}_m\) the variety of Borel subalgebras of \(\gl_m\).
Let \(H \subseteq \gl_m\) be the subalgebra of upper triangular matrices.
The variety of Borel subalgebras of \(\gl_m\) is \(\mathcal{B} = \{gHg^{-1} : g \in \GL_m\}\).
Thus, the Springer fiber at \(X \in \mathcal{N}\) is 
\[S_X = \{gHg^{-1} : X \in gHg^{-1}\}.\]
\begin{definition}
    A \emph{flag} \((V_i)\) of \(\mathbb{C}^m\) is a sequence of subspaces
    \[0 = V_0 \subseteq V_1 \subseteq \cdots \subseteq V_m = \mathbb{C}^m,\]
    where \(\dim V_i = i\).
\end{definition}
We say that \(X \in \gl_m\) \emph{preserves} a flag \((V_i)\) if \(\forall i. \; XV_i \subseteq V_i\).
Note that \(X \in \mathcal{N}\) preserves a flag \((V_i)\) if and only if \(\forall i. \; XV_i \subseteq V_{i - 1}\).
\par The simplest flag is the \emph{standard flag} \((E_i)\), where \(E_i := \langle e_1, ..., e_i \rangle\).
Note that the group \(H\) is exactly the subset of \(\gl_m\) which preserves \(E_i\).
\par We can think of \(\mathcal{B}\) as the set of flags of \(\mathbb{C}^m\), via the correspondence 
\[gHg^{-1} \leftrightarrow (gE_i).\]
Note that \(X\) preserves \((gE_i)\) if and only if \(X \in gHg^{-1}\).
Thus, we may write the Springer fiber at \(X \in \mathcal{N}\) in terms of flags, as 
\[S_X = \{(gE_i) : X \in gHg^{-1}\} = \{(V_i) : \forall i. \; XV_i \subseteq V_{i - 1}\}.\]

@@

\section{Future work}
The paper (TODO: cite) describes a bijection between the Steinberg variety (which is blah) and pairs of standard Young tableaux (each having shape blah).
The variety \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\) looks something like the Steinberg variety.
Maybe there is a similar correspondence, between elements of \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\) and pairs of standard Young tableaux: one having the shape of the Jordan form of \(X\), one having the shape of the Jordan form of \(A_{X, a, b}\).
\par Separately, this paper would benefit from an introductory section where we motivate the study of \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).

>>>>

\usepackage{hyperref}
\usepackage[capitalize,nameinlink,noabbrev]{cleveref} % to emulate \autoref style
% \usepackage{fullpage}

@@

\author{
UROP+ Final Paper, Summer 2024\\
Owen Conoly\\
Mentor: Haoshuo Fu\\
Project Suggested by: Haoshuo Fu
}
\date{}
\title{Irreducible Components and Dimension of the Springer Fiber of a Hook-Type Slodowy Slice}

% email to andre when done
\begin{document}

\maketitle
\begin{abstract}
    with more words than symbols, what is springer fiber of hook-type slodowy slice.
    we show that its irreducible components have the same dimensions.
    and use this to study another variety.
\end{abstract}

\tableofcontents

\section{Introduction}
Let \(\mathfrak{g} = \gl_m \times \gl_{m + n}\), with \(n > 0\).
% The nilpotent cone in \(\mathfrak{g}\) is \(\mathcal{N} = \mathcal{N}_m \times \mathcal{N}_{m + n}\), where \(\mathcal{N}_m\) and \(\mathcal{N}_{m + n}\) are the nilpotent cones in \(\gl_m\) and \(\gl_{m + n}\) respectively.
% Let \(\mathcal{B}_m\) and \(\mathcal{B}_{m + n}\) be the varieties of Borel subalgebras of \(\gl_m\) and \(\gl_n\) respectively.
% The variety of Borel subalgebras of \(\mathcal{N}\) is \(\mathcal{B} = \mathcal{B}_m \times \mathcal{B}_{m + n}\).
% Let
% \[\widetilde{\mathcal{N}} := \{(\mathfrak{b}, n) \in \mathcal{B} \times \mathcal{N} : n \in \mathfrak{b}\} = \]
% \[\{(\mathfrak{b}_m, \mathfrak{b}_{m + n}, n_m, n_{m + n}) \in \mathcal{B}_m \times \mathcal{B}_{m + n} \times \mathcal{N}_m \times \mathcal{N}_{m + n} : n_m \in \mathfrak{b}_m, n_{m + n} \in \mathfrak{b}_{m + n}\}.\]
Let \(\pi : \widetilde{\mathcal{N}} \to \mathcal{N} \hookrightarrow \mathfrak{g}\) be the Springer resolution.
\par Let \((e, h, f)\) be the principal \(\spl_2\)-triple in \(\gl_n\).
We have an embedding \(\gl_n \hookrightarrow \gl_m \times \gl_n \hookrightarrow \gl_{m + n}\).
Let \((E, H, F)\) be the \(\spl_2\)-triple in \(\gl_{m + n}\) which is the image of \((e, h, f)\) via this embedding.
Let \(S\) be the Slodowy slice \(E + \mathfrak{z}_{\gl_m}(F)\).
We have a map \(\gl_{m + n} \to \gl_m\) given by coordinate projection.
Restricting to \(S\), we obtain \(f_1 : S \to \gl_m\).
(It turns out that this map is surjective.)
Then, we obtain \(p_1 : S \to \mathfrak{g}\) by \(x \mapsto (x, f_1(x))\).
Taking the map \(p_1 : S \to \mathfrak{g}\) and the Springer resolution \(\widetilde{\mathcal{N}} \to \mathfrak{g}\), we obtain a fibred product \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).
\par In this paper we study \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).
% The first few sections will be finding what it looks like (in particular, what are the elements of \(S\) which project to nilpotent elements of \(\mathfrak{g}\)), and then we will find the irreducible components of \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).
As a step towards studying \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\), we consider the map \(S \times_\mathfrak{g} \widetilde{\mathcal{N}} \to \gl_m\) given by taking the map to \(\mathfrak{g}\) and then projecting to \(\gl_m\).
The main section of this paper will study the fibers of this map.
Given \(X \in \mathcal{N}_m \subseteq \gl_m\), we call the fiber at \(X\) the \emph{\(n\)-Slodowy-slice Springer fiber at \(X\)}.
\par Before talking about future work, say what is the main result of the paper.

\subsection{Future Work}
\par irreducble components have module structure of group algebra of Sn.
has applications to representation theory.
write about this here.
The paper (TODO: cite) describes a bijection between the Steinberg variety (which is blah) and pairs of standard Young tableaux (each having shape blah).
The variety \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\) looks something like the Steinberg variety.
Maybe there is a similar correspondence, between elements of \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\) and pairs of standard Young tableaux: one having the shape of the Jordan form of \(X\), one having the shape of the Jordan form of \(A_{X, a, b}\).

\subsection{Overview}
\par Section 2 reviews some preliminary material.
Section 3 finds the unique (up to similarity) principal \(\spl_2\)-triple in \(\gl_n\).
Section 4 embeds this \(\spl_2\)-triple into \(\gl_{m + n}\) as described above.
We compute the Slodowy slice and end up with a nice description of \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).
\par Section 5 discusses how to reduce the problem of finding the irreducible components of an \(n\)-Slodowy-slice Springer fiber at \(X \in \mathcal{N}_m\) to an easier problem.
Section 6 solves the easier problem.
Section 7 finds the irreducible components of an \(n\)-Slodowy-slice Springer fiber.
Section 8 applies the results of section 7 to find the irreducible components of \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).
Section 9 discusses an interesting direction for future work.
Finally, section 10 proves some linear algebra lemmas that were used in the paper.

@@

\subsection{Springer fibers in \texorpdfstring{\(\gl_m\)}{gl\_m}}
Now we let \(\mathcal{N}_m\) be the set of nilpotent elements in \(\gl_m\), and \(\mathcal{B}_m\) the variety of Borel subalgebras of \(\gl_m\).
Let \(H \subseteq \gl_m\) be the subalgebra of upper triangular matrices.
The variety of Borel subalgebras of \(\gl_m\) is \(\mathcal{B} = \{gHg^{-1} : g \in \GL_m\}\).
Thus, the Springer fiber at \(X \in \mathcal{N}\) is 
\[\mathcal{F}_X = \{gHg^{-1} : X \in gHg^{-1}\}.\]
% ^ use that notation everywhere for springer fiber
\begin{definition}
    A \emph{flag} \((V_i)\) of \(\mathbb{C}^m\) is a sequence of subspaces
    \[0 = V_0 \subseteq V_1 \subseteq \cdots \subseteq V_m = \mathbb{C}^m,\]
    where \(\dim V_i = i\).
\end{definition}
We say that \(X \in \gl_m\) \emph{preserves} a flag \((V_i)\) if \(\forall i. \; XV_i \subseteq V_i\).
Note that \(X \in \mathcal{N}\) preserves a flag \((V_i)\) if and only if \(\forall i. \; XV_i \subseteq V_{i - 1}\).
\par The simplest flag is the \emph{standard flag} \((E_i)\), where \(E_i := \langle e_1, ..., e_i \rangle\).
Note that the group \(H\) is exactly the subset of \(\gl_m\) which preserves \(E_i\).
\par We can think of \(\mathcal{B}\) as the set of flags of \(\mathbb{C}^m\), via the correspondence 
\[gHg^{-1} \leftrightarrow (gE_i).\]
Note that \(X\) preserves \((gE_i)\) if and only if \(X \in gHg^{-1}\).
Thus, we may write the Springer fiber at \(X \in \mathcal{N}\) in terms of flags, as 
\[S_X = \{(gE_i) : X \in gHg^{-1}\} = \{(V_i) : \forall i. \; XV_i \subseteq V_{i - 1}\}.\]

@@

% \section{Future work}
% \par Separately, this paper would benefit from an introductory section where we motivate the study of \(S \times_\mathfrak{g} \widetilde{\mathcal{N}}\).
